Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-05 writeup/intro.tex
More editing

Our main contributions in this paper are as follows. We relate the
many previously proposed isolation and consistency models to the goal
high availability, which guarantees a response from every data replica
in the presence of arbitrary partitions between them.  We classify
which among the wide array of models are achievable with high
availability, denoting them as {\em Highly Available Transactions}
(HATs). In doing so, we demonstrate that although many implementations
of HAT semantics are not highly available, this is an artifact of the
implementations rather than an inherent property of the semantics. Our
investigation shows that, besides serializability, Snapshot Isolation
and Repeatable Read isolation are not HAT-compliant, while most other
isolation levels are achievable with high availability. We also
demonstrate that many weak replica consistency models from distributed
systems are both HAT-compliant and simultaneously achievable with
several ACID properties.

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we show is
unavailable. However, Read Committed isolation, transactional
atomicity (``Monotonic Atomic View'' isolation), and many other consistency
models from database and distributed systems are achievable via
algorithms that rely on multi-versioning and limited client-side
caching. For several guarantees, such as causal consistency with
phantom prevention and ANSI Repeatable Read, we consider a modified
form of high availability in which clients ``stick to'' (i.e., have
affinity with) at least one server---a property which is often
implicit in the distributed systems literature~\cite{herlihy-art,
  eiger, cac} but which requires explicit consideration in a
client-server replicated database context. This sticky availability is
widely employed~\cite{eiger, vogels-defs} but is a less restrictive
model (and therefore more easily achievable) than traditional high
availability.

@@

Finally, we recognize that the large variety of ACID isolation levels
and distributed consistency models (and therefore those in our
taxonomization) can be confusing; the subtle distinctions between
models can appear to be of primarily academic concern. Accordingly, we
offer the following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

>>>>

Our main contributions in this paper are as follows. We relate the
many previously proposed isolation and consistency models to the goal
of high availability, which guarantees a response from each non-failed
data replica in the presence of arbitrary partitions between them.  We
classify which among the wide array of models are achievable with high
availability, denoting them as {\em Highly Available Transactions}
(HATs). In doing so, we demonstrate that although many implementations
of HAT semantics are not highly available, this is an artifact of the
implementations rather than an inherent property of the semantics. Our
investigation shows that, besides serializability, Snapshot Isolation
and Repeatable Read isolation are not HAT-compliant, while most other
isolation levels are achievable with high availability. We also
demonstrate that many weak replica consistency models from distributed
systems are both HAT-compliant and simultaneously achievable with
several ACID properties.

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we show is
unavailable. However, Read Committed isolation, transactional
atomicity (Section~\ref{sec:ta}), and many other consistency models
from database and distributed systems are achievable via algorithms
that rely on multi-versioning and limited client-side caching. For
several guarantees, such as causal consistency with phantom prevention
and ANSI Repeatable Read, we consider a modified form of high
availability in which clients ``stick to'' (i.e., have affinity with)
at least one server---a property which is often implicit in the
distributed systems literature~\cite{herlihy-art, eiger, cac} but
which requires explicit consideration in a client-server replicated
database context. This sticky availability is widely
employed~\cite{eiger, vogels-defs} but is a less restrictive model
(and therefore more easily achievable) than traditional high
availability.

@@

Finally, we recognize that the large variety of ACID isolation levels
and distributed consistency models (and therefore those in our
taxonomization) can be confusing; the subtle distinctions between
models may appear to be of academic concern. Accordingly, we offer the
following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

### 2013-10-05 writeup/model.tex
More editing

\subsection{Replica Availability}

@@

If a client wishes to execute a transaction that performs operations
on multiple data items, then the client must be able to contact and
receive a response from at least one replica for each data item. We
say that a transaction has \textbf{replica availability} if it can
contact at least one replica for every item it attempts to
access. This may result in ``lower availability'' than a
non-transactional requirement. Additionally, given the possibility of
system-initiated aborts, we need to ensure useful forward progress: a
system can trivially guarantee clients a response by always aborting
all transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}. We cannot
guarantee that every transaction will commit---transactions may choose
to abort themselves---but we need to make sure that the system will
not indefinitely abort transactions on its own volition. We call a
transaction abort due to a transaction's own choosing (e.g., as an
operation of the transaction itself or due to a would-be violation of a declared integrity constraint) an \textit{internal abort} and an
abort due to system implementation or operation an \textit{external
  abort}. We say that a system provides \textbf{transactional
  availability} if, given replica availability for every data item in
a transaction, the transaction eventually commits (possibly after
multiple client retries) or internally aborts~\cite{hat-hotos}.

>>>>

@@

We say that a transaction has \textbf{replica availability} if it can
contact at least one replica for every item it attempts to access;
this may result in ``lower availability'' than a non-transactional
availability requirement (e.g., single-item
availability). Additionally, given the possibility of system-initiated
aborts, we need to ensure useful forward progress: a system can
trivially guarantee clients a response by always aborting all
transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}.

A system cannot guarantee that every transaction will
commit---transactions may choose to abort themselves---but we need to
make sure that the system will not indefinitely abort transactions on
its own volition. We call a transaction abort due to a transaction's
own choosing (e.g., as an operation of the transaction itself or due
to a would-be violation of a declared integrity constraint) an
\textit{internal abort} and an abort due to system implementation or
operation an \textit{external abort}. We say that a system provides
\textbf{transactional availability} if, given replica availability for
every data item in a transaction, the transaction eventually commits
(possibly after multiple client retries) or internally
aborts~\cite{hat-hotos}. A system provides \textbf{sticky
  transactional availability} if, given sticky availability, a
transaction eventually commits or internally aborts.
